Details of specific functions

### 2021-07-28 inst/gitbook/03_specific-tutorials.Rmd
typos up

There are many ways of measuring disparity!
In brief, disparity is a summary metric that will represent an aspect of an ordinated space (e.g. a MDS, PCA, PCO, PCoA).
For example, one can look at ellipsoid hyper-volume of the ordinated space (Donohue _et al._ 2013), the sum and the product of the ranges and variances (Wills _et al._ 1994) or the median position of the elements relative to their centroid (Wills _et al._ 1994). 
Of course, there are many more examples of metrics one can use for describing some aspect of the ordinated space, with some performing better than other ones at particular descriptive tasks, and some being more generalist.
Check out this paper on selecting the best metric for your specific question in [Ecology and Evolution](https://onlinelibrary.wiley.com/doi/full/10.1002/ece3.6452).
<!-- biorXiv preprint -->
You can also use the  [`moms` shiny app](https://tguillerme.shinyapps.io/moms/) to test which metric captures which aspect of traitspace occupancy regarding your specific space and your specific question.

@@

Level | Name | Description | Source | 
------|------|---------------------------------------------------|--------|
2 | `ancestral.dist` | The distance between an element and its ancestor | `dispRity` |
2 | `angles` | The angle of main variation of each dimensions | `dispRity` |
2 | `centroids`<sup>1</sup> | The distance between each element and the centroid of the ordinated space | `dispRity` |
1 | `convhull.surface` | The surface of the convex hull formed by all the elements | [`geometry`](https://cran.r-project.org/web/packages/geometry/index.html)`::convhulln$area` |
1 | `convhull.volume` | The volume of the convex hull formed by all the elements | [`geometry`](https://cran.r-project.org/web/packages/geometry/index.html)`::convhulln$vol` |
2 | `deviations` | The minimal distance between each element and a hyperplane | `dispRity` |
1 | `diagonal` | The longest distance in the ordinated space (like the diagonal in two dimensions) | `dispRity` |
2 | `displacements` | The ratio between the distance from a reference and the distance from the centroid | `dispRity` |
1 | `edge.length.tree` | The edge lengths of the elements on a tree | `ape` |
1 | `ellipse.volume`<sup>1</sup> | The volume of the ellipsoid of the space | Donohue _et al._ (2013) |
1 | `func.div` | The functional divergence (the ratio of deviation from the centroid) | `dispRity` (similar to [`FD`](https://cran.r-project.org/web/packages/FD/index.html)`::dbFD$FDiv` but without abundance)|
1 | `func.eve` | The functional evenness (the minimal spanning tree distances evenness) | `dispRity` (similar to [`FD`](https://cran.r-project.org/web/packages/FD/index.html)`::dbFD$FEve` but without abundance)|
1 | `group.dist` | The distance between two groups | `dispRity` |
1 | `mode.val` | The modal value | `dispRity` |
1 | `n.ball.volume` | The hyper-spherical (*n*-ball) volume | `dispRity` |
2 | `neighbours` | The distance to specific neighbours (e.g. the nearest neighbours - by default) | `dispRity` |
2 | `pairwise.dist` | The pairwise distances between elements | [`vegan`](https://cran.r-project.org/web/packages/vegan/index.html)`::vegist` |
2 | `point.dist` | The distance between one group and the point of another group | `dispRity` |
2 | `projections.tree` | The `projections` metric but where the vector can be based on a tree | `dispRity` |
2 | `projections` | The distance *on* (projection) or *from* (rejection) an arbitrary vector | `dispRity` |
2 | `quantiles` | The *n*th quantile range per axis | `dispRity` |
2 | `radius` | The radius of each dimensions | `dispRity` |
2 | `ranges` | The range of each dimension | `dispRity` |
2 | `span.tree.length` | The minimal spanning tree length | [`ape`](https://cran.r-project.org/web/packages/vegan/index.html) (but see also [`vegan`](https://cran.r-project.org/web/packages/vegan/index.html)`::spantree`) |
2 | `variances` | The variance of each dimension | `dispRity` |

>>>>

There are many ways of measuring disparity!
In brief, disparity is a summary metric that will represent an aspect of an ordinated space (e.g. a MDS, PCA, PCO, PCoA).
For example, one can look at ellipsoid hyper-volume of the ordinated space (Donohue _et al._ 2013), the sum and the product of the ranges and variances (Wills _et al._ 1994) or the median position of the elements relative to their centroid (Wills _et al._ 1994). 
Of course, there are many more examples of metrics one can use for describing some aspect of the ordinated space, with some performing better than other ones at particular descriptive tasks, and some being more generalist.
Check out this paper on selecting the best metric for your specific question in [Ecology and Evolution](https://onlinelibrary.wiley.com/doi/full/10.1002/ece3.6452).
You can also use the  [`moms` shiny app](https://tguillerme.shinyapps.io/moms/) to test which metric captures which aspect of traitspace occupancy regarding your specific space and your specific question.

@@

Level | Name | Description | Source | 
------|------|---------------------------------------------------|--------|
2 | `ancestral.dist` | The distance between an element and its ancestor | `dispRity` |
2 | `angles` | The angle of main variation of each dimensions | `dispRity` |
2 | `centroids`<sup>1</sup> | The distance between each element and the centroid of the ordinated space | `dispRity` |
1 | `convhull.surface` | The surface of the convex hull formed by all the elements | [`geometry`](https://cran.r-project.org/web/packages/geometry/index.html)`::convhulln$area` |
1 | `convhull.volume` | The volume of the convex hull formed by all the elements | [`geometry`](https://cran.r-project.org/web/packages/geometry/index.html)`::convhulln$vol` |
2 | `deviations` | The minimal distance between each element and a hyperplane | `dispRity` |
1 | `diagonal` | The longest distance in the ordinated space (like the diagonal in two dimensions) | `dispRity` |
2 | `displacements` | The ratio between the distance from a reference and the distance from the centroid | `dispRity` |
1 | `edge.length.tree` | The edge lengths of the elements on a tree | `ape` |
1 | `ellipse.volume`<sup>1</sup> | The volume of the ellipsoid of the space | Donohue _et al._ (2013) |
1 | `func.div` | The functional divergence (the ratio of deviation from the centroid) | `dispRity` (similar to [`FD`](https://cran.r-project.org/web/packages/FD/index.html)`::dbFD$FDiv` but without abundance)|
1 | `func.eve` | The functional evenness (the minimal spanning tree distances evenness) | `dispRity` (similar to [`FD`](https://cran.r-project.org/web/packages/FD/index.html)`::dbFD$FEve` but without abundance)|
1 | `group.dist` | The distance between two groups | `dispRity` |
1 | `mode.val` | The modal value | `dispRity` |
1 | `n.ball.volume` | The hyper-spherical (*n*-ball) volume | `dispRity` |
2 | `neighbours` | The distance to specific neighbours (e.g. the nearest neighbours - by default) | `dispRity` |
2 | `pairwise.dist` | The pairwise distances between elements | [`vegan`](https://cran.r-project.org/web/packages/vegan/index.html)`::vegist` |
2 | `point.dist` | The distance between one group and the point of another group | `dispRity` |
2 | `projections.tree` | The `projections` metric but where the vector can be based on a tree | `dispRity` |
2 | `projections` | The distance *on* (projection) or *from* (rejection) an arbitrary vector | `dispRity` |
2 | `quantiles` | The *n*th quantile range per axis | `dispRity` |
2 | `radius` | The radius of each dimensions | `dispRity` |
2 | `ranges` | The range of each dimension | `dispRity` |
2 | `span.tree.length` | The minimal spanning tree length | [`vegan`](https://cran.r-project.org/web/packages/vegan/index.html)`::spantree` |
2 | `variances` | The variance of each dimension | `dispRity` |
